A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-10-19 doc/large_strain.tex
show deformed meshes
more loading BC more straight forward, also update the text

In this section, we apply the proposed matrix-free operator algorithms for finite-strain hyperelastic materials as well as the geometric multigrid preconditioner to a benchmark problem from \cite{Miehe2007}. The coarse meshes for the 2D and 3D problems are illustrated in Figure \ref{fig:miehe}. The 2D material consists of a square (depicted in blue), a hole and two spherical inclusions (depicted in red). The size of the square domain is $10^{-3} \rm{mm}$.
The matrix material is taken to have Poisson's ratio $0.3$ and shear modulus $\mu=0.4225 \times 10^6\, \rm{N/mm^2}$. The inclusion is taken to be $100$ times stiffer.
The 3D material is obtained by extrusion of the 2D geometry into the third dimension.
Note that, although a relatively coarse mesh is used at the coarsest multigrid level, the geometry of the inclusions is captured accurately by manifold descriptions of the boundaries and interfaces.
Both domains are fully fixed along the bottom surface and a distributed load is applied at the top in the $(1,0)$ or $(1,1,0)$ direction for the 2D and the 3D problems, respectively.
The distributed load of $0.5\,\rm{N}$ is applied in 5 steps.
With respect to the nonlinear solver, the displacement tolerance of the $\mathcal{\ell}_2$ norm for the Newton update is taken to be $10^{-5}$ whereas the relative and absolute tolerances for the residual forces is $10^{-8}$. The relative convergence criteria for the linear solver is $10^{-6}$.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \centering
    \includegraphics[width=\textwidth]{gmg_2d.png}
    % \caption{2D multigrid mesh with 2 global refinements}
  \end{subfigure}
  \caption{2D multigrid mesh after two global refinements distributed into three MPI processes. The color indicates the process owning the element.}%
  \label{fig:miehe_gmg}
\end{figure}
%
Figure \ref{fig:miehe_gmg} illustrates the hierarchy of multigrid meshes for the 2D problem after two global refinements.

@@

The benchmark calculations are performed for 2D and 3D problems with various combination of polynomial degrees and number of global refinements, as is stated in Tables \ref{tab:input_parameters_2d} and \ref{tab:input_parameters_3d}.
Shared memory parallelization using Intel Threading Building Blocks was disabled for this study.
%

We are interested in the following metrics:
\begin{itemize}
\item the wall-clock time and memory requirement per DoF for matrix-vector multiplication;
\item the average number of CG iterations throughout the entire simulation (i.e. each Newton-Raphson iteration and each loading step);
\item the wall-clock time per DoF for the CG solver;
\end{itemize}
%
The results of the matrix-free approach with a geometric-multigrid preconditioner are compared to the matrix-based approach in conjunction with the algebraic multigrid (AMG) preconditioner using the package ML of the Trilinos \cite{Heroux2005} library, version 12.12.1.
The aggregation threshold for the AMG preconditioner is taken as $10^{-4}$.
Computations are done on a single node of two clusters:
A node on the ``Emmy'' cluster at RRZE, FAU has two Xeon 2660v2 ``Ivy Bridge'' chips (10 cores per chip + SMT) running at 2.2 GHz with 25 MB shared cache per chip and 64 GB of RAM. Intel's compiler version 18.03 with flags ``-O3 -march=native'' and Intel MPI version 18.03 are used.
The ``IWR'' results were obtained on a single machine with eight Xeon E7-8870 ``Sandy Bridge'' chips (10 cores per chip  + SMT) running at 2.4 GHz with 30 MB shared cache per chip. For the simulation, GCC version 8.1.0 with flags ``-O3 -march=native'' and OpenMPI version 3.0.0 were used.

>>>>

In this section, we apply the proposed matrix-free operator algorithms for finite-strain hyperelastic materials as well as the geometric multigrid preconditioner to a benchmark problem from \cite{Miehe2007}. The coarse meshes for the 2D and 3D problems are illustrated in Figure \ref{fig:miehe}. The 2D material consists of a square (depicted in blue), a hole and two spherical inclusions (depicted in red). The size of the square domain is $10^{-3} \rm{mm}$.
The matrix material is taken to have Poisson's ratio $0.3$ and shear modulus $\mu=0.4225 \times 10^6\, \rm{N/mm^2}$. The inclusion is taken to be $100$ times stiffer.
The 3D material is obtained by extrusion of the 2D geometry into the third dimension.
Note that, although a relatively coarse mesh is used at the coarsest multigrid level, the geometry of the inclusions is captured accurately by manifold descriptions of the boundaries and interfaces.
Both domains are fully fixed along the bottom surface and a distributed load is applied at the top in the $(1,0)$ or $(1,1,0)$ direction for the 2D and the 3D problems, respectively.
The force density $12.5 \times 10^3 \rm{N/mm^2}$ or $12.5 \times 10^3 \rm{N/mm^3}$ is applied in 5 steps for the 2D and the 3D problems, respectively.
With respect to the nonlinear solver, the displacement tolerance of the $\mathcal{\ell}_2$ norm for the Newton update is taken to be $10^{-5}$ whereas the relative and absolute tolerances for the residual forces is $10^{-8}$. The relative convergence criteria for the linear solver is $10^{-6}$.
Shared memory parallelization using Intel Threading Building Blocks was disabled for this study.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \centering
    \includegraphics[width=\textwidth]{gmg_2d.png}
    % \caption{2D multigrid mesh with 2 global refinements}
  \end{subfigure}
  \caption{2D multigrid mesh after two global refinements distributed into three MPI processes. The color indicates the process owning the element.}%
  \label{fig:miehe_gmg}
\end{figure}
%
Figure \ref{fig:miehe_gmg} illustrates the hierarchy of multigrid meshes for the 2D problem after two global refinements.
Figure \ref{fig:miehe_deformed} shows deformed meshes at the final loading step with quadratic elements and two global mesh refinements.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{deformed.png}
      \caption{2D deformed mesh}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{deformed_3d.png}
    \caption{3D deformed mesh}
  \end{subfigure}
  \caption{Deformed meshes at the final loading step with quadratic elements and two global mesh refinements.}%
  \label{fig:miehe_deformed}
\end{figure}

@@

The benchmark calculations are performed for 2D and 3D problems with various combination of polynomial degrees and number of global refinements, as is stated in Tables \ref{tab:input_parameters_2d} and \ref{tab:input_parameters_3d}.
%
We are interested in the following metrics:
\begin{itemize}
\item the wall-clock time and memory requirement per DoF for matrix-vector multiplication;
\item the average number of CG iterations throughout the entire simulation (i.e. each Newton-Raphson iteration and each loading step);
\item the wall-clock time per DoF for the CG solver;
\end{itemize}
%
The results of the matrix-free approach with a geometric-multigrid preconditioner are compared to the matrix-based approach in conjunction with the algebraic multigrid (AMG) preconditioner using the package ML of the Trilinos \cite{Heroux2005} library, version 12.12.1.
The aggregation threshold for the AMG preconditioner is taken as $10^{-4}$.
Computations are done on a single node of two clusters:
A node on the ``Emmy'' cluster at RRZE, FAU has two Xeon 2660v2 ``Ivy Bridge'' chips (10 cores per chip + SMT) running at 2.2 GHz with 25 MB shared cache per chip and 64 GB of RAM. Intel's compiler version 18.03 with flags ``-O3 -march=native'' and Intel MPI version 18.03 are used.
The ``IWR'' results were obtained on a single machine with eight Xeon E7-8870 ``Sandy Bridge'' chips (10 cores per chip  + SMT) running at 2.4 GHz with 30 MB shared cache per chip. For the simulation, GCC version 8.1.0 with flags ``-O3 -march=native'' and OpenMPI version 3.0.0 were used.
